A spiking neural integrator model of the adaptive control of action by the medial prefrontal cortex

### 2013-11-14 paper/rebuttal.tex

\begin{quoting}
  In relation to the previous point, it seems quite ad hoc and looks
  like a strong simplification to hand tune the model so that "reward
  delivery [...] resets the system to a starting state" while "an
  error trial [drives it] to a low state". Shouldn't different
  reactions of the model to correct versus incorrect trials emerge
  through learning (such as in reinforcement learning-based models of
  ACC; e.g. Alexander and Brown, 2011; Khamassi et al., 2013)? Or
  shouldn't they at least produce symmetrical effects on the model
  that it should exploit to produce different reactions? Isn't this a
  strong help provided a priori to the model to make it react in the
  expected way? Maybe I am wrong to think that this is a
  simplification. Maybe this is equivalent to the priors put in the
  cited previous computational models. But this should be at least
  discussed by the authors. For instance, doesn't such simplification
  prevent generalizations of the model to tasks involving a sequence
  of rewards and/or punishments (e.g. Tanaka et al., 2004)?
\end{quoting}

This is an important point that identifies
the central difference between
models created with the NEF
and many other computational modeling approaches,
including reinforcement learning-based approaches.
What Reviewer 1 calls hand tuning and simplification
is essentially our dynamical systems hypothesis
of how to solve the reaction time task
with a control system operating in continuous time.
While it may seem a simplification
for reward to drive the system to the origin
and error to drive it to a low part of the state space,
these reflect the operations
that are possible (and, indeed, natural) in dynamical systems.
We use these operations to implement a dynamical system
that performs the task
as we hypothesize the rodent performs the task.
We can then implement that dynamical system
in a network of spiking neural populations
through the principles of the NEF.
This spiking neural network models
the neural circuit of a rodent
that has already learned the task;
going from random connectivity to this network
is a separate branch of research.
We believe that providing the end-point of learning
is a great help to modelers
aiming to learn this kind of neural circuit
(and have shown an example of this
in other work; e.g., \citenoparens{MacNeil2011}).

In other words, the NEF provides three biologically plausible
principles for network construction
through which we can produce a neural circuit.
We have presented a way in which those principles
can be used to generate a network
that performs the simple RT task.
We have attempted to
emphasize these consideration
in various parts of the revised discussion,
most notably Sections 5.1 and 5.2.

\begin{quoting}
  Some simulation results are too briefly presented -- sometimes
  making it difficult to understand for the reader -- and some are not
  analyzed with enough deepness. On page 10, the results displayed in
  figure 4 should be better explained in the main text. A few lines
  below, in contrast to what the authors wrote, it is not clear for
  the reader from these trajectories that the time of the cue can be
  predicted. During correct trials, the cue has already occurred to
  make ACC in the right starting position for the next trial. Please
  explain what you mean here.
\end{quoting}

Simulations of the dynamical system are expanded
and simplified, as discussed below.
We have included two new figures (4 and 5)
that aim to help with understanding
how the simulations work under different sets of parameters.
More detailed equations throughout Sections 2.2 and 2.3
are also included in the revised manuscript.

\begin{quoting}
  The authors should show, either on figs 4 and 5 or on supplementary
  figures, correct, premature and late trajectories starting from the
  lower-left starting position (following an error).
\end{quoting}

Figure 4 in the revised manuscript now presents trajectories
from many different initial conditions.
It also shows how the result of the last trial
determines the initial conditions
of the next trial.

\begin{quoting}
  Can the states be decoded in the same manner when starting from such
  a different position?
\end{quoting}

The state can be decoded,
but the cue cannot be accurately predicted
unless the system starts at the origin
under normal system behavior.
This is why cue prediction does not occur
in post-error trials;
the system starts in the lower part
of the state space instead of the origin.
This is shown and expanded on in Figure 4
and Section 3.1.

\begin{quoting}
  On page 11, the authors show that the double integrator model has
  the same PCs as the experimental data. Could the authors also show
  whether the cue-responsive model also produces similar PCs as
  experimental data?
\end{quoting}

The cue-responsive network does not contain any integrative
populations, so it is highly unlikely that
the PCs would be the same as the experimental data.
However, there is an open question
(which was raised by Reviewer 2)
as to whether other network structures
would result in the same principal
components as the double integrator network.
In order to address this question,
we have added Figure 8 in the revised manuscript,
which shows that
a single integrator does not produce
similar PCs as experimental data.

\begin{quoting}
  In Narayanan and Laubach (2009), premature and late cases are
  combined in a single PC1 and PC2 couple of curves for "post-error".
  Do the PCs of the original data (from Narayanan and Laubach 2009)
  still look like the model's curves when premature and late cases are
  separated?
\end{quoting}

In \citetext{Narayanan2009}, the PC analysis
was only done using correct trials.
Error trials were not included in the analysis.
PCA done on premature trials
deviate from those based on correct trials
due to the rats responding early in the foreperiod.
PCA on late trials do not differ much
from those on correct trials,
but lack modulation during the normal period
when rats release the lever
at the end of the foreperiod.

\citetext{Narayanan2008},
which was mistakenly not cited in the previous
version of the paper, reported effects
for premature and late responses,
and showed that single units
(not population activity)
tracked outcomes during
the intertrial interval.

Importantly, while neural activity
tracked outcomes during the intertrial interval,
that activity did not differentiate
between the type of error that occurred
\cite{Narayanan2008}.
Therefore, we combine post-premature
and post-late trials when analyzing
the experimental data
in order to increase
the number of trials being analyzed,
and therefore increasing
the statistical power of the analysis
For the simulated data,
we analyze the two separately
in order to show that the model also
does not differentiate between
the two cases;
since we can simulate as many
trials as we want,
we do not need to combine
the types of error trials
in order to increase statistical power.

\begin{quoting}
  Could the authors explain how late trials can occur in the adaptive
  control model nearly as often as the other model while the former
  anticipates and responds faster than the latter?
\end{quoting}

Late trials can occur in exactly the same way
as in the cue-responding model post-error,
as the adaptive control model takes
a cue-responding strategy in that case.
Like in the cue-responding case,
late trials often correspond to non-responses,
in which the release should have occurred,
but release sensitive neurons were
not active for long enough to actually
effect the full lever press.
These errors occur at approximately
the same frequency in the cue-responding
and adaptive control models.

\begin{quoting}
  Can the model also reproduce activities during the second version of
  the task in Narayanan and Laubach 2009 where cues are presented only
  on half of the trials?
\end{quoting}

Because the cue is predicted by the double integrator network,
the model can accomplish the task even if the cue
is not actually present.
The model would not change
or produce different PCs in this case,
so we did not discuss the second version of the task
in order to keep the focus of the paper clear.

\begin{quoting}
  Do the simulation results really demonstrate that the ACC cannot
  inhibit the motor command? (Page 9). If ACC is x1+x2, then what
  brain structure correspond to H+T+P/R units? If it is also ACC, the
  authors need to first validate their ability to reproduce recorded
  ACC activity. If not, then how does the model prove that ACC is not
  inhibiting motor responses? A convincing demonstration would have
  been to simulate the two alternative hypotheses and show that one
  fails to reproduce some experimental evidence. Otherwise, the
  authors should not sell this part of the work (page 9) as a strong
  demonstration, but rather just say that with this ACC model, it is
  not necessary to exert inhibition on motor cells to work.
\end{quoting}

We have removed sections
discussing the motor inhibition hypothesis
from the paper.
We thank Reviewer 1 for pointing out
the weaknesses of this argument;
we believe that the revised discussion
proposes more concrete predictions
that come from the model.

\begin{quoting}
  The prediction extracted from the model on page 12 ("our model
  predicts that fewer premature releases would occur") appears like a
  trivial prediction. It seems to the reader that there is no need of
  a computational model to predict this. Either expand on this
  prediction or give arguments to convince the reader that this is not
  a trivial prediction.
\end{quoting}

We agree that this prediction is trivial,
and have removed it from the revised manuscript.

\begin{quoting}
  The model should not only be evaluated in its ability to reproduce
  neurophysiological recordings, but also in its ability to reproduce
  rat behavior simultaneously recorded. For instance, on page 10, it
  is stated that "if the previous trial was an error, the double
  integrator will not be following the trajectory that predicts the
  time of the cue, and therefore will not cause a lever release." Is
  it what rats do in Narayanan and Laubach (2009)? Please quantify
  such behavior and other behaviors of the model and compare with rat
  behavioral data.
\end{quoting}

Unfortunately, it is impossible to determine if rodents
are predicting the time of the cue given their behavior alone;
we could estimate this by analyzing
the reaction times on correct trials
following an error,
but due to amount of variability in reaction times,
there are not enough instances of post-error correct trials
to definitively state that reaction times are slower
on these trials compared to post-correct correct trials.

We provide a summary comparing
the behavioral statistics of
the spiking cue-responding and adaptive control networks
to the experimental subjects in Figure 10.
We believe this is the most appropriate evaluation
given the data available.

\begin{quoting}
  The manuscript would be strengthened if the authors could provide a
  series of testable experimental predictions that would allow to
  further test/refute the proposed computational model. In particular,
  what further experiments are needed to decide between the two
  mentioned alternative hypotheses about ACC inhibiting motor
  responses or this inhibition being elsewhere?
\end{quoting}

We have added Section 5.3 listing several predictions.
Since we do not discuss the motor inhibition hypothesis,
we do not propose an experiment relating to this hypothesis.
However, we do make predictions about
how neurons might respond in tasks
with different temporal dynamics,
and discuss experiments that can be done
to confirm or refute our predictions.

\begin{quoting}
  The discussion about the necessity to differentiate between
  adaptability due to recurrent activity and adaptability due to
  synaptic weight changes is very interesting and should be further
  elaborated and extended. In particular, previous ACC models rely on
  reinforcement learning mechanisms that would induce adaptability due
  to synaptic weight changes (Alexander and Brown, 2011; Khamassi et
  al., 2013). Here the model produces adaptability due to recurrent
  activity. It would be useful to discuss whether some of the results
  produced by these previous models may be also explained by recurrent
  activity. If yes, could the authors think of an experimental
  protocol that would enable to decide which of the alternative models
  is right?
\end{quoting}

We agree that differentiating
between adaptability
through recurrent activity
and through synaptic weight changes
is an important issue
that we think will become increasingly important
as more complex mPFC models are proposed.
We have added Section 5.2 to discuss
this differentiation.

In the revised discussion, we have proposed two
new experiments that could allow for
direct comparisons between neural data
and the simulated networks.
First, one could record as rats are initially trained
to perform the task and compare
the dynamics of the mPFC (and motor cortex)
with a double integrator network.
A challenge for the theory side would be
to evaluate various types of RL algorithms
that might enable a match
between theory and data.
Second, one could train rats to perform
a variant of the simple RT task
using multiple foreperiods,
and test the rats in sessions with
interleaved and blocked foreperiods.
If the mPFC neurons
and double integrator neurons track
the expected timing of action,
as we predict, then there would be
shifts in the temporal expectations
following each switch in block
(e.g., from 0.5 to 1.5 second foreperiods
over blocks of 30 trials).
To our knowledge,
no group has carried out such studies.
However, they could allow for
direct comparisons between
brain activity and theory.

\begin{quoting}
  The discussion about possible extension of the model to account for
  data in the Stroop task should be discussed in relation to previous
  models. What would such extension of the authors' model add to
  previous simulations with the model of Dehaene and Changeux (1998)?
  Globally, the discussion should be extended to thoroughly address
  each raised issue.
\end{quoting}

We have removed the possible extension of the model
to the Stroop task and instead highlighted
the model's specificity to the simple RT task.
However, we have also highlighted in Section 5.2
that the process of creating
dynamical system for performing the simple RT task
could also be done for many other tasks
and implemented using the same methods.

\section*{Reviewer 2}

We would like to thank Reviewer 2
for noting the interesting point that
the double integrator model and the published
mPFC recordings share the same
population statistics, and for inquiring
about some details in the neural circuit model.
We believe that we have been able to
address all of his or her concerns,
and, as above for Reviewer 1,
the manuscript is much improved
thanks to his or her suggestions.
In what follows,
we detail our responses to each comment.

\begin{quoting}
  I think the most compelling result in the paper is in figure 6 --
  i.e. that the network has the same principal components as the
  neural data. One question I have is to what extent this result is
  unique to a double integrator or could other networks produce these
  kind of PCs? My feeling is you have to have something like a double
  integrator but I would like to see some control simulations that
  show that. So, for example, would a single integrator network have
  those PCs? Would a random network with the same inputs have those
  PCs?
\end{quoting}

This is an interesting point,
especially when using principal component analysis
on filtered and Z-scored spike trains.
To address this issue, we added
principal component analysis results
from a single integrator network (Figure 8),
and show that it does not have
the same PCs as the double integrator
or the neural recordings.

\begin{quoting}
  I'd like to see slightly more in section 2.3.2. In particular what
  are the exact update equations for the network? In particular what
  is the exact form of $G$ what are the values of $J_{bias}$ etc ... I
  realize these details are in the referenced paper and book but I
  think it is appropriate to have them in an abridged form here.
\end{quoting}

The update equation for $G$ has been included in Section 2.4
(which was formerly Section 2.3.2),
and a table of the parameters used
(that discusses how $J_{bias}$ is determined)
has been included as Table 1.

\begin{quoting}
  I might have missed this - but how many neurons are in the model?
  You say you sample 174 but I'm not sure how many are in the actual
  model?
\end{quoting}

Thanks for catching this issue.
The simulated networks had 1200 neurons;
this has been stated in the revised text
and in Figure 3.

\begin{quoting}
  Regarding the fact that the behavior of the model is slightly too
  good (not enough errors) -- you mention adding noise to the model
  might solve this problem. I think it would be worth reporting
  whether it does in practice. Neural integrators can be fickle beasts
  so I don't think this is a given.
\end{quoting}

Our original mention of noise was in reference
to the spike density functions plotted
in (now) Figures 7 and 8.
While the principal components matched closely,
the spike densities revealed that
the simulated data is much cleaner
than the experimentally recorded data.
As Reviewer 2 suggests, we added noise
($J_\eta$, now included in equations
in Section 2.4)
and recreated Figure 7 (formerly Figure 6).
As can be seen in the revised manuscript,
the simulated spike densities now very closely resemble
the experimental spike densities,
while still maintaining a very close
principal component fit.

In regards to the model performing too well behaviorally,
this is not an issue of the noisiness of the system,
but on other modeling decisions.
In particular, we could have varied $\beta$
in order to obtain more premature errors,
which would match the behavior
of experimental subjects more closely.
However, rerunning the simulations
that are summarized in Figure 10
takes several days.
Since having a perfect fit
to the experimental performance
is not necessary to match other quantities
like reaction times,
we chose an acceptable value for $\beta$
based on direct simulations
of the dynamical system and
did not vary it while simulating the spiking implementation.

\begin{quoting}
  I think it would be nice if you could generate some testable
  predictions of the model. For example, can you say anything about
  the character of neurons that contribute to each of the principal
  components ... is there any difference between them in the model in
  terms of connectivity (are PC2 neurons more highly connected than
  PC1 neurons) and sign (are PC1 neurons more likely to be
  inhibitory). What about microstimulation or inactivation? Similarly
  for the holding, trigger and press release neurons - how would you
  know if you were recording them? As I understand it they might not
  look like figure 8 because that's the decoded signal not firing
  rate. These are just suggestions and I will leave it to the authors
  to try and find something - but the nice thing about a model is you
  can interrogate it and I think it would improve the paper to put
  down some firm predictions that can be tested.
\end{quoting}

Thanks for this suggestion,
which was also raised by Reviewer 1.
Testable predictions have been added in Section 5.3.

\begin{quoting}
  Will the Nengo scripts be available through J Neuroscience? I'm not
  quite sure what the policy is for this kind of material but it would
  be nice if it were available with the article as the github may
  change over time ...
\end{quoting}

A link to the files used
for the submitted version of the manuscript
is provided in the revised manuscript.
This link will always point to the
version of those files
used to generate the revised manuscript
even if changes are subsequently
made to those files.
Additionally, we have also included
at that link a complete copy
of the Nengo simulator
used to produce the model,
in case changes to Nengo also
affect model behavior.
With this, any computing environment
with Java should be able to
run the simulations,
and a computing environment
with Python and several open source libraries
(listed with the analysis scripts)
should be able to run
the analysis and generate
the plots shown in the paper.

\bibliographystyle{namedplus}
\bibliography{jneurosci2013}

\end{document}
